**OptiFit: a fast method for fitting amplicon sequences to existing OTUs**

### 2021-01-19 paper/manuscript.Rmd
initial outline

**(observation format - max 1200 words, 2 figures, 25 ref)**

@@

@@

## Results

@@

## Materials and Methods
(data availability)

>>>>

**observation format - max 1200 words, 2 figures, 25 ref**

@@

- CRC is one of the most common cancers and a leading cause of cancer related death
- There is evidence that the microbiome has a role in CRC development/progression and could be useful for biomarker detection and diagnostics.
- Begum et al (mBio 2020) recently demonstrated effective application of machine learning (ML) to microbiome based classification problems and developed a framework for applying ML practices in a more reproducible way (mikropml).
- A common question when applying ML methods to microbiome data is which method and taxonomic level should be use.
- This analysis utilizes the reproducible framework developed by Begum et al to quantify which ML method and taxonomic level produce the best performing classifier for CRC data.

@@

## Results
- Of the five ML methods tested, Random forest was consistently the top performer (supplemental figure of all models?) at most taxonomic levels.
  - RF might be more appropriate anyways since its more suitable for zero inflated data? (need to look into literature)
- Within the RF model, the highest AUCs were observed for family, genus, and otu level data with no significant difference between the three. (Figure 1)

@@

## Materials and Methods
- data from prior study {baxter}
- mikropml package
- pvalues as previously described {begum}

### 2021-01-19 paper/manuscript.tex
initial outline

@@

@@

(data availability)

>>>>

\begin{itemize}
\tightlist
\item
  CRC is one of the most common cancers and a leading cause of cancer
  related death
\item
  There is evidence that the microbiome has a role in CRC
  development/progression and could be useful for biomarker detection
  and diagnostics.
\item
  Begum et al (mBio 2020) recently demonstrated effective application of
  machine learning (ML) to microbiome based classification problems and
  developed a framework for applying ML practices in a more reproducible
  way (mikropml).
\item
  A common question when applying ML methods to microbiome data is which
  method and taxonomic level should be use.
\item
  This analysis utilizes the reproducible framework developed by Begum
  et al to quantify which ML method and taxonomic level produce the best
  performing classifier for CRC data.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Of the five ML methods tested, Random forest was consistently the top
  performer (supplemental figure of all models?) at most taxonomic
  levels.

  \begin{itemize}
  \tightlist
  \item
    RF might be more appropriate anyways since its more suitable for
    zero inflated data? (need to look into literature)
  \end{itemize}
\item
  Within the RF model, the highest AUCs were observed for family, genus,
  and otu level data with no significant difference between the three.
  (Figure 1)
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  data from prior study \{baxter\}
\item
  mikropml package
\item
  pvalues as previously described \{begum\}
\end{itemize}
